A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-03-04 analysis/paper/paper.Rmd
using HDR instead of quantiles for shoreline dates

Where x is [...] In Figure \@ref(fig:shoredate) this formula is used to shoreline date the same sites from where this relationship was derived. Following from having defined the distance between intersecting sea- and site polygons as zero during simulations, the sites were dated using the mean elevation of the site polygons to allow for some deviance in elevation. The result of comparing the synchroneity of the dates using the method presented by @parnell2008 is given in Figure \@ref(fig:shoredate)B. Here, 100,000 age samples drawn from each ^14^C-date was subtracted from 100,000 age samples drawn from the corresponding shoreline date. The resulting range between the 2.5th and 97.5th percentile were then checked to see if they cross zero, in which case the dates are considered synchronous. The result of this procedure indicates that the shoreline date correspond with the radiocarbon dates for 44 out of 49 sites (`r round(44/49, 2)*100`%). The deviation from the shoreline date at the two Løvås sites and Gunnarsrød 5 is to be expected based on the simulation results (see above), but not at Langangen Vestgård 1 and Nordby 52. For these two sites the zero age difference can be seen to fall just outside the 95% interval (Figure \@ref(fig:shoredate)B), and in the simulation of the distance to the shoreline this was simulated to be 0m across virtually all of the 1000 simulation runs for both sites (see supplementary material). If one accepts that sensitivity in the method leads to the failure to accurately date these two sites, and that the failed dating of the two Løsvås sites and Gunnarsrød 5 are instead due to extraneous factors and not the method itself, one could alternatively see the success rate as defined by 44 out of 46 sites (`r round(44/46, 2)*100`%) having been correctly dated.

>>>>

Where x is [...] In Figure \@ref(fig:shoredate) this formula is used to shoreline date the same sites from where this relationship was derived. Following from having defined the distance between intersecting sea- and site polygons as zero during simulations, the sites were dated using the mean elevation of the site polygons to allow for some deviance in elevation. The result of comparing the synchroneity of the dates using the method presented by @parnell2008 is given in Figure \@ref(fig:shoredate)B. Here, 100,000 age samples drawn from each shoreline date was subtracted from 100,000 age samples drawn from the corresponding ^14^C-date. The resulting range of the 95% highest density region (HDR, @ ) was then checked to see if it crosses zero, in which case the dates are considered synchronous. The result of this procedure indicates that the shoreline date correspond with the radiocarbon dates for 46 out of 49 sites (`r round(46/49, 2)*100`%). The deviation from the shoreline date at Gunnarsrød 5 is to be expected based on the simulation results (see above), but not at Langemyr and Kvastad A2. For these two sites the zero age difference can be seen to fall just outside the 95% interval (Figure \@ref(fig:shoredate)B), and in the simulation of the distance to the shoreline this was simulated to be 0m across virtually all of the 1000 simulation runs for both sites (see supplementary material). If one accepts that sensitivity in the method leads to the failure to accurately date these two sites, and that the failed dating of the two Løsvås sites and Gunnarsrød 5 are instead due to extraneous factors and not the method itself, one could alternatively see the success rate as defined by 44 out of 46 sites (`r round(44/46, 2)*100`%) having been correctly dated.
